The effects of varying depth in cosmic shear surveys

### 2019-12-16 Heydenreich_et_al.tex
accepted version

   \date{Received 21 October 2019; accepted XXX}

@@

%
%-------------------------------------------------------------------
\section{Introduction}
The discovery of cosmic shear has provided us with a new and powerful cosmological tool to empirically test the standard model of cosmology and to determine its parameters. Contrary to the analysis of the cosmic microwave background \citep[CMB, e.g.~by][]{2018arXiv180706209P}, cosmic shear is more sensitive to the properties of the low-redshift large-scale structure and thus provides an excellent consistency check for the standard model. Current cosmic shear surveys are particularly sensitive to the parameter $S_8=\sigma_8 \sqrt{\Omega_{\rm m}/0.3}$, where $\sigma_8$ characterizes the normalization of the matter power spectrum and $\Omega_{\rm m}$ is the matter density parameter. Constraints on $S_8$ from the three current major cosmic shear results are all consistent with the CMB analysis by \citet{2018arXiv180706209P}. It is interesting to note, however, that they all favour values that are slightly lower than the Planck constraints of $S_8 = 0.830 \pm 0.013$. \citet{2018arXiv180909148H} report $S_8 = 0.800^{+0.029}_{-0.028}$ from an analysis of the Subaru Hyper Suprime-Cam survey, \citet[][hereafter H18]{2018arXiv181206076H} obtained $0.737_{-0.036}^{+0.040}$ from KiDS+VIKING data, and \citet{2018PhRvD..98d3528T} constrain $S_8=0.782\pm 0.027$ using the Dark Energy Survey (DES).  Combined analyses of DES and KiDS data \citep{Joudaki:2019, asgari/etal:2019} results in a $\sim 3 \sigma$ tension with the CMB value for $S_8$. If this tension is not the manifestation of an unaccounted systematic effect, in either the cosmic shear surveys \citep{Mandelbaum:2019} or the Planck mission \citep{2016ApJ...818..132A}, it certainly merits attention. It could be interpreted as a sign of new physics with examples like massive neutrinos \citep{2014PhRvL.112e1303B}, time-varying dark energy or modified gravity \citep{2016A&A...594A..14P} and coupling within the dark sector \citep{Kumar:2019}.  It could also, however, be a simple statistical coincidence.

@@

For current cosmic shear surveys, the estimated systematic error is becoming comparable in size to the statistical error, implying that for next-generation surveys, a significant reduction of systematic errors is necessary. With surveys like the Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic:2008} and Euclid \citep{Laureijs:2011} soon to start, systematic effects in gravitational lensing have received a large amount of attention \citep[see][and references therein]{Mandelbaum:2019}.

In this paper we focus on systematic effects induced by variation in survey depth that is so far unaccounted for in cosmic shear analyses \citep{Vale:2004}.   For a fixed exposure time survey, varying atmospheric conditions, dithering strategies and galactic extinction all contribute to an inhomogeneous limiting magnitude as a function of sky position.   In order to assess the impact of variable depth for current and future surveys we build an analytical model for the effect based on the survey specifications of the Kilo-Degree Survey \citep[KiDS,][]{Kuijken:2015}.  To first order, the depth variation in KiDS can be modelled by a piece-wise constant depth function, which varies between each $1\,\text{deg}^2$ square pointing.   KiDS object detection is defined in the $r$-band as these images were chosen to be significantly deeper in comparison to the other optical and near-infra red filters.   We therefore quantify survey depth with the limiting $r$-band magnitude, as defined in \citet{2017A&A...604A.134D}.  We defer the study of multi-band variable depth and its impact on photometric redshift accuracy for future work.

{\bf This work is complementary to the analysis of \citet{Guzik:2005} who investigate the effect of a general, position-dependent multiplicative shear bias on the shear power spectrum.  In principle the varying depth of the source galaxy sample that we study here could be recast as a varying effective shear bias.  The inclusion of an inhomogeneous distribution of source galaxies has also been explored using mock catalogs of the Subaru Hyper Suprime-Cam Survey \citep{Shirasaki:2019}, with a focus on resulting estimation of the cosmic shear covariance matrix.

}

@@

In Sect. \ref{sec:analytic_models} we will introduce two simple toy models to understand this effect and analyze the impact on the cosmic shear power spectrum. In Sect. \ref{sec:xipm_semianalytic} we will estimate the effect on the shear correlation functions $\xi_\pm$ using a semi-analytic model. We will present our results in Sect. \ref{sec:results}. In Sect. \ref{sec:discussion} we will discuss our results and comment on the impact of our used simplifications. In the appendices we present the full derivation of our model for finite field surveys.
%App.~A is reserved for detailed calculations that cannot be done in the main paper, in App.~B we introduce way to apply our model to a given survey footprint. App.~C is reserved for figures and tables that cannot be displayed in the main body. 
We will assume the standard weak gravitational lensing formalism, a summary of which can be found in \citet{2001PhR...340..291B}.

%-------------------------------------------------------------------
\section{Simple, analytic toy models}
\label{sec:analytic_models}
For our first analysis we assume that all the matter between the sources and observer is concentrated in a single lens plane of distance $D_{\rm d}$ from the observer. If we now distribute sources at varying distances $D_{\rm s}$, two effects become apparent: Firstly, the lensing efficiency $D_{\rm{ds}}/D_{\rm s}$ varies, where $D_{\rm{ds}}$ is the distance between the lens plane and the respective source. Secondly, and more importantly, for a more distant source \emph{more} matter is concentrated between the source and the observer, leading to a stronger shear signal.

Assuming that the depth, and thus the source redshift population, only varies between pointings of the camera, an observer will measure a shear signal that is modified by a step-like depth-function, $\gammao(\b\theta)=W(\b \theta)\gamma(\b\theta)$, where $W$ is proportional to the mean of the lensing efficiency $\Dds/\Ds$ of one pointing and $\gamma$ denotes the shear that this pointing would experience if it were of the average depth.
We can parametrize $W$ as $W(\bm{\theta}) = 1+w(\bm{\theta})$. This implies that $\langle w(\bm{\theta})\rangle=0$ holds, where $\langle \cdot\rangle$ denotes the average over all pointings.
\subsection{Modelling the power spectrum}
\label{sec:modelling_power_spectrum}
In our first model we describe the impact of varying depth on the power spectrum, following the simplifications described above.
% Following \citet{1993ApJ...404..441K}, we can simulate the shear $\gamma$ from the convergence by convolution with the kernel \begin{equation}
%\mathcal{D}(\bm{\theta}) = \frac{\theta_2^2-\theta_1^2-2i\theta_1\theta_2}{\pi |\bm{\theta}|^4}\, .
%\end{equation}
%
%Similarly, an observer that measures the shear $\gamma$ can reconstruct the convergence $\kappa$ by convolution with the kernel $\mathcal{D}^*$, where ${}^*$ denotes the complex conjugate. Assuming that the optical depth, and thus the source redshift populations, vary between pointing, an observer will measure a shear-signal that is modified by a depth-function $W(\bm{\theta}) = 1-w(\bm{\theta})$, where $\langle w(\bm{\theta})\rangle=0$ holds. The reconstruction of the observed convergence $\kappa^{\rm obs}$ will thus not yield the original convergence. In particular there is no reason why the convolution with a complex kernel should have a nonzero imaginary part. This allows us to separate our received signal into E- and B-modes following \begin{equation}
% \kappa = \kappa^E+i\kappa^B\, ,
%\end{equation}
%where $B$-Modes cannot be generated by a lensing signal and are thus a tracer of remaining systematics.
%2002A&A...389..729S SRC
%
In accordance with the definition of the shear power spectrum \begin{equation}
\la \hat{\gamma}(\b\ell)\,\hat{\gamma}^*(\b\ell')\ra = (2\pi)^2\delta(\b\ell-\b\ell')P(|\b\ell|) \, ,
\label{eq:original_power_spectrum}
where $\hat{\gamma}$ denotes the Fourier transform of $\gamma$, we define the observed power spectrum via \begin{equation}
P^{\text{obs}}(\b\ell) \equiv \frac{1}{(2\pi)^2}\int \d^2 \ell'\,  \la \gammaoh(\b \ell)\, \gammaoh {}^*(\b \ell')\ra \, .
Note that due to the depth-function, both the assumptions of homogeneity and isotropy break down, which means that we can neither assume isotropy in the power spectrum, nor can we assume that $\la \gammaoh(\b \ell) \gammaoh {}^*(\b \ell')\ra$ vanishes for $\b\ell\neq\b\ell'$. This estimator provides a natural extension to the definition of the regular power spectrum and, in the case of a homogeneous depth distribution, reduces back to the original estimator.
To model a constant depth on each individual pointing, $\b \alpha$, we can choose random variables, $w_{\b \alpha}$, that only need to satisfy $\langle w_{\b \alpha}\rangle=0$. As we assume an infinite number of pointings, $\b\alpha$ can assume any two-dimensional integer value $\mathbb{Z}^2$ and we can parametrize $w(\b\theta)$ as 
w(\b \theta) = \sum_{\b \alpha \in \mathbb{Z}^2} w_{\b \alpha} \Xi(\b \theta-L\b \alpha)\, ,
with the box-function
\Xi(\b \theta) = \begin{cases}
\end{cases},
\label{eq:defweightf}
where $L$ is the sidelength of one pointing.
Following the calculations in App.~\ref{sec:calc of PS}, we derive 
P^{\text{obs}}(\b \ell)  =  P(\b \ell) + \la w^2\ra \int\frac{\text{d}^2\b \ell'}{(2\pi)^2}\,\hat{\Xi}(\b \ell-\b \ell')\, P(\b \ell')\, .
Here we have denoted $\la w^2\ra \equiv \la w_{\b \alpha}^2\ra$ as the dispersion of the depth-function, since the statistical properties of this function do not depend on the pointing $\b\alpha$. The Fourier transform of the box function, $\hat{\Xi}$, is a 2-dimensional sinc-function (see \ref{sec:calc of PS}).
The observed power spectrum, $P^{\text{obs}}$, is thus composed of the original power spectrum $P(\ell)$ from Eq. \eqref{eq:original_power_spectrum}, plus a convolution of the power spectrum with a sinc-function, scaling with the variance of the function $w(\b\theta)$.

@@

\subsection{Modelling the shear correlation functions}
\label{sec:xipm_analytic}
More convenient measures to infer cosmological information from observational data are the shear correlation functions $\xi_\pm$, which are defined as \begin{equation}
\xi_\pm(\theta) = \la \gamma_{\rm t}\gamma_{\rm t}\ra(\theta) \pm \la \gamma_\times\gamma_\times\ra(\theta) \, .
\end{equation}
Here, $\gamma_{\rm{t}}$ and $\gamma_\times$ denote the tangential- and cross-component of the shear for a galaxy pair with respect to their relative orientation \citep[see][]{2002A&A...396....1S}.
The shear correlation functions are the prime estimators to quantify a cosmic-shear signal, since it is simple to include a weighting of the shear measurements into the correlation functions and, contrary to the power spectrum, one does not have to worry about the shape of the survey footprint or masked regions, or model the noise contribution. 
For this analysis we will follow the assumption that a deeper pointing shows a stronger shear signal $\gamma^{\rm{obs}}(\b\theta) = W(\b\theta)\gamma(\b\theta)$ as described above. This assumption implies that a higher redshift just increases the amplitude of the shear signal, but as can clearly be seen by inspecting shear correlation functions of different redshift distributions, the change of the signal is extremely scale-dependent and not just a multiplication with a constant factor. In other words, not just the average shear changes as a function of redshift, but also its entire two-point statistics. However, this should serve as a reasonable first approximation for small variations in mean source redshift. Additionally, we assume that a greater depth does not only lead to a stronger average shear, but also to a higher galaxy number density, implying a correlation between those two quantities.

Let $N^i(\b \theta),N^j(\b\theta)$ be the average weighted number of galaxies\footnote{Instead of using the actual number of galaxies, we take the effective number density, as defined in \citet{Kuijken:2015}, scaled by the respective survey area. Due to this we account for different weighting of galaxies in the shear correlation functions as well as in the average redshift distribution.} per pointing in redshift bins $i$ and $j$ and let $W^i(\b \theta),W^j(\b\theta)$ be the weighting of average shear. The observed correlation functions $\xi^{ij,\text{obs}}_\pm(\theta)$ now change from uniform depth, $\xi_\pm^{ij,\rm{uni}}(\theta)$, via %\citep[see][]{2002A&A...396....1S}
\begin{align}
\xi^{ij,\text{obs}}_\pm(\theta) = & \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_{\rm t}(\bth)\gamma^{j,\rm{obs}}_{\rm t}(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \\
 & \pm \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_\times(\bth)\gamma^{j,\rm{obs}}_\times(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \nonumber\\
 = & \frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i(\bth)N^j(\bthp)\ra} \xi_{\pm}^{ij,\rm{uni}}(\theta) \, ,
 \label{eq:xipmblub1}
 \end{align}
 where the average $\la\cdot\ra$ represents both an ensemble average as well as an average over the position $\bth$.
 Assuming that the depth of different pointings is uncorrelated, the only important property of a galaxy pair is whether or not they lie in the same pointing. We want to denote the probability that a random galaxy pair of separation $\theta$ lies in the same pointing by $E(\theta)$. This function is depicted in Fig. \ref{fig:eoftheta_lin}, and an analytic expression is derived in App. \ref{sec:model_e}.

@@

 Ignoring that the depth is correlated on scales of one pointing (here at $\theta \leq \sqrt{2}^\circ$) is equivalent to setting $E(\theta)\equiv 0$. Note that there is still a correlation between $N$ and $W$ for the same galaxy.

@@

 Performing the same calculations as above, this yields a relation between the correlation function of uniform depth, $\xi_\pm^{ij,\rm{uni}}$, and the one that is usually modelled, $\xi_\pm^{ij}$:

@@

When an observer now calculates the model correlation functions $\xi^{ij}_\pm$ without accounting for varying depth between pointings, the ratio between modelled and observed correlation functions becomes: \begin{align}

@@

It is interesting to note that $\xi^{ij}_\pm = \xi_\pm^{ij,\rm{obs}}$ holds wherever $E(\theta)=0$, so we expect the observed and the modelled correlation functions to be equivalent on scales where the depth is uncorrelated. One thing left to determine is how to define the weight-function $W(\b\theta)$. For this, we will refer the reader to the beginning of Sec.~\ref{sec:results}.

@@

\section{A semi-analytic model}
\label{sec:xipm_semianalytic}
The previously derived analytic model describes how varying depth between pointings modifies the correlation function due to the correlation between number density and the average redshift of source galaxies. While this model serves as an intuitive first approximation, it completely ignores any effects from the large scale structure (LSS) between the closest and the most distant galaxy. Therefore, we do not expect this model to yield accurate, quantitative results for cosmic shear surveys.

@@

An inspection of KiDS-data showed that the redshift distribution of sources is highly correlated with the limiting magnitude in the $r$-band. We thus chose to separate the survey into 10 quantiles, sorted by $r$-band depth, i.e. if a pointing had a shallower depth than 90\% of the other pointings, it would belong to the first quantile, and so on. For each quantile $m$ and each tomographic redshift bin $i$ we can extract a weighted number of galaxies $N^i_m$ and a source redshift distribution $p^i_m(z)$ following the direct spectroscopic calibration method of H18. In Fig.~\ref{fig:nz_of_meanz} the average redshift and weighted number of galaxies are plotted for each quantile of each redshift bin, whereas a selection of source redshift distributions is depicted in Fig.~\ref{fig:nofz_quantiles}. A table of the limiting magnitudes for each quantile can be found in Tab.~\ref{tab:maglim}.
%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{images/cov_nz_meanz.pdf}
\caption{Weighted number of galaxies $N$ and average redshift $\la z\ra$ in the KiDS+VIKING-450 survey \citep[KV450,][]{Wright:2018} in pointings of different depth for each of the five tomographic bins used in H18. Each colour corresponds to one redshift bin of H18. A single point represents one quantile of the respective redshift bin, where the fainter points denote pointings of shallower depth.}
\label{fig:nz_of_meanz}
\end{figure}

\begin{figure}
\centering
\includegraphics[trim = {1cm 0 1.2cm 0.1cm}, clip, width=\linewidth]{images/dist_z_all.pdf}
\caption{The source redshift distributions $p^i_m(z)$ for a selection of very shallow pointings (blue), average pointings (yellow) and very deep pointings (green). The percentage points in the legend denote to which quantile a pointing belongs, when all are ordered by their depth.}
\label{fig:nofz_quantiles}
\end{figure}
% Asserting an arbitrary cosmological model, we can transform the source redshift distributions $p^i_m(z)$ into comoving distance probability distributions $\mathcal{L}^i_m(\chi)$.

>>>>

   \date{Received 21 October 2019; accepted 02 December 2019}

@@

%
%-------------------------------------------------------------------
\section{Introduction}
The discovery of cosmic shear has provided us with a new and powerful cosmological tool to empirically test the standard model of cosmology and to determine its parameters. Contrary to the analysis of the cosmic microwave background \citep[CMB, e.g.,~by][]{2018arXiv180706209P}, cosmic shear is more sensitive to the properties of the low-redshift large-scale structure and, thus, provides an excellent consistency check for the standard model. Current cosmic shear surveys are particularly sensitive to the parameter $S_8=\sigma_8 \sqrt{\Omega_{\rm m}/0.3}$, where $\sigma_8$ characterizes the normalization of the matter power spectrum and $\Omega_{\rm m}$ is the matter density parameter. Constraints on $S_8$ from the three current major cosmic shear results are all consistent with the CMB analysis by \citet{2018arXiv180706209P}. It is interesting to note, however, that they all favor values that are slightly lower than the Planck constraints of $S_8 = 0.830 \pm 0.013$. \citet{2018arXiv180909148H} report $S_8 = 0.800^{+0.029}_{-0.028}$ from an analysis of the Subaru Hyper Suprime-Cam survey, \citet[][hereafter H18]{2018arXiv181206076H} obtained $0.737_{-0.036}^{+0.040}$ from KiDS+VIKING data, and \citet{2018PhRvD..98d3528T} constrain $S_8=0.782\pm 0.027$ using the Dark Energy Survey (DES).  Combined analyses of DES and KiDS data \citep{Joudaki:2019, asgari/etal:2019} result in a $\sim 3 \sigma$ tension with the CMB value for $S_8$. If this tension is not the manifestation of an unaccounted systematic effect, in either the cosmic shear surveys \citep{Mandelbaum:2019} or the Planck mission \citep{2016ApJ...818..132A}, it certainly merits attention. It could be interpreted as a sign of new physics exemplified by massive neutrinos \citep{2014PhRvL.112e1303B}, time-varying dark energy, or modified gravity \citep{2016A&A...594A..14P} and coupling within the dark sector \citep{Kumar:2019}.  It could also, however, prove to be a simple statistical coincidence.

@@

For current cosmic shear surveys, the estimated systematic error is becoming comparable in magnitude to the statistical error, implying that for next-generation surveys, a significant reduction of systematic errors is necessary. With surveys like the Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic:2008} and Euclid \citep{Laureijs:2011} soon to begin, systematic effects in gravitational lensing have received a large amount of attention \citep[see][and references therein]{Mandelbaum:2019}.

In this paper, we focus on systematic effects induced by variation in survey depth that is so far unaccounted for in cosmic shear analyses \citep{Vale:2004}.   For a survey with a fixed exposure time, varying atmospheric conditions, dithering strategies and galactic extinction all contribute to an inhomogeneous limiting magnitude as a function of sky position.   In order to assess the impact of variable depth for current and future surveys, we build an analytical model for the effect based on the survey specifications of the Kilo-Degree Survey \citep[KiDS,][]{Kuijken:2015}.  To the first order, the depth variation in KiDS can be modeled by a piece-wise constant depth function, which varies between each $1\,\text{deg}^2$ square pointing.   KiDS object detection is defined in the $r$-band as these images were chosen to be significantly deeper in comparison to the other optical and near-infra red filters.   We, therefore, quantify survey depth with the limiting $r$-band magnitude, as defined in \citet{2017A&A...604A.134D}.  We defer the study of multi-band variable depth and its impact on photometric redshift accuracy for a future work.

This work is complementary to the analysis of \citet{Guzik:2005}, who investigate the effect of a general, position-dependent multiplicative shear bias on the shear power spectrum.  In principle, the varying depth of the source galaxy sample that we study here could be recast as a varying effective shear bias.  The inclusion of an inhomogeneous distribution of source galaxies has also been explored using mock catalogs of the Subaru Hyper Suprime-Cam Survey \citep{Shirasaki:2019}, with a focus on resulting estimation of the cosmic shear covariance matrix.

@@

In Sect. \ref{sec:analytic_models}, we will introduce two simple toy models to understand this effect and analyze the impact on the cosmic shear power spectrum. In Sect. \ref{sec:xipm_semianalytic}, we will estimate the effect on the shear correlation functions $\xi_\pm$ using a semi-analytic model. We will present our results in Sect. \ref{sec:results}. In Sect. \ref{sec:discussion}, we will discuss our results and comment on the impact of our used simplifications. In the appendices, we present the full derivation of our model for finite field surveys.
%App.~A is reserved for detailed calculations that cannot be done in the main paper, in App.~B we introduce way to apply our model to a given survey footprint. App.~C is reserved for figures and tables that cannot be displayed in the main body. 
We assume the standard weak gravitational lensing formalism, a summary of which can be found in \citet{2001PhR...340..291B}.

%-------------------------------------------------------------------
\section{Simple, analytic toy models}
\label{sec:analytic_models}
For our first analysis, we assume that all the matter between the sources and observer is concentrated in a single lens plane of distance $D_{\rm d}$ from the observer. If we then distribute sources at varying distances $D_{\rm s}$, two effects become apparent: firstly, the lensing efficiency $D_{\rm{ds}}/D_{\rm s}$ varies, where $D_{\rm{ds}}$ is the distance between the lens plane and the respective source. Secondly, and more importantly, for a more distant source, more matter is concentrated between the source and the observer, leading to a stronger shear signal.

Assuming that the depth and, thus, the source redshift population, only varies between pointings of the camera, an observer will measure a shear signal that is modified by a step-like depth-function, $\gammao(\b\theta)=W(\b \theta)\gamma(\b\theta)$, where $W$ is proportional to the mean of the lensing efficiency $\Dds/\Ds$ of one pointing and $\gamma$ denotes the shear that this pointing would experience if it were of the average depth.
We can parametrize $W$ as $W(\bm{\theta}) = 1+w(\bm{\theta})$. This implies that $\langle w(\bm{\theta})\rangle=0$ holds, where $\langle \cdot\rangle$ denotes the average over all pointings.

\subsection{Modelling the power spectrum}
\label{sec:modelling_power_spectrum}
In our first model, we describe the impact of varying depth on the power spectrum, following the simplifications described above.
% Following \citet{1993ApJ...404..441K}, we can simulate the shear $\gamma$ from the convergence by convolution with the kernel \begin{equation}
%\mathcal{D}(\bm{\theta}) = \frac{\theta_2^2-\theta_1^2-2i\theta_1\theta_2}{\pi |\bm{\theta}|^4}\, .
%\end{equation}
%
%Similarly, an observer that measures the shear $\gamma$ can reconstruct the convergence $\kappa$ by convolution with the kernel $\mathcal{D}^*$, where ${}^*$ denotes the complex conjugate. Assuming that the optical depth, and thus the source redshift populations, vary between pointing, an observer will measure a shear-signal that is modified by a depth-function $W(\bm{\theta}) = 1-w(\bm{\theta})$, where $\langle w(\bm{\theta})\rangle=0$ holds. The reconstruction of the observed convergence $\kappa^{\rm obs}$ will thus not yield the original convergence. In particular there is no reason why the convolution with a complex kernel should have a nonzero imaginary part. This allows us to separate our received signal into E- and B-modes following \begin{equation}
% \kappa = \kappa^E+i\kappa^B\, ,
%\end{equation}
%where $B$-Modes cannot be generated by a lensing signal and are thus a tracer of remaining systematics.
%2002A&A...389..729S SRC
%
In accordance with the definition of the shear power spectrum \begin{equation}
\la \hat{\gamma}(\b\ell)\,\hat{\gamma}^*(\b\ell')\ra = (2\pi)^2\delta(\b\ell-\b\ell')P(|\b\ell|) \, ,
\label{eq:original_power_spectrum}
\end{equation}
where $\hat{\gamma}$ denotes the Fourier transform of $\gamma$, we define the observed power spectrum via \begin{equation}
P^{\text{obs}}(\b\ell) \equiv \frac{1}{(2\pi)^2}\int \d^2 \ell'\,  \la \gammaoh(\b \ell)\, \gammaoh {}^*(\b \ell')\ra \, .
\end{equation}
We note that due to the depth-function, both the assumptions of homogeneity and isotropy break down, which means that we can neither assume isotropy in the power spectrum, nor can we assume that $\la \gammaoh(\b \ell) \gammaoh {}^*(\b \ell')\ra$ vanishes for $\b\ell\neq\b\ell'$. This estimator provides a natural extension to the definition of the regular power spectrum and, in the case of a homogeneous depth distribution, is reduced back to the original estimator.
To model a constant depth on each individual pointing, $\b \alpha$, we can choose random variables, $w_{\b \alpha}$, that only need to satisfy $\langle w_{\b \alpha}\rangle=0$. As we assume an infinite number of pointings, $\b\alpha$ can assume any two-dimensional integer value $\mathbb{Z}^2$ and we can parametrize $w(\b\theta)$ as 
\begin{equation}
w(\b \theta) = \sum_{\b \alpha \in \mathbb{Z}^2} w_{\b \alpha} \Xi(\b \theta-L\b \alpha)\, ,
\end{equation}
with the box-function
\begin{equation}
\Xi(\b \theta) = \begin{cases}
1 \qquad \b \theta\in \left[-\frac{L}{2},\frac{L}{2}\right]^2 \\
0 \qquad \text{else}
\end{cases},
\label{eq:defweightf}
\end{equation}
where $L$ is the sidelength of one pointing.
Following the calculations in App.~\ref{sec:calc of PS}, we derive 
\begin{equation}
P^{\text{obs}}(\b \ell)  =  P(\b \ell) + \la w^2\ra \int\frac{\text{d}^2\b \ell'}{(2\pi)^2}\,\hat{\Xi}(\b \ell-\b \ell')\, P(\b \ell')\, .
\end{equation}
Here we have denoted $\la w^2\ra \equiv \la w_{\b \alpha}^2\ra$ as the dispersion of the depth-function, since the statistical properties of this function do not depend on the pointing $\b\alpha$. The Fourier transform of the box function, $\hat{\Xi}$, is a 2-dimensional sinc-function (see \ref{sec:calc of PS}).
The observed power spectrum, $P^{\text{obs}}$, is thus composed of the original power spectrum $P(\ell)$ from Eq. \eqref{eq:original_power_spectrum}, plus a convolution of the power spectrum with a sinc-function, scaling with the variance of the function $w(\b\theta)$.

@@

\subsection{Modeling the shear correlation functions}
\label{sec:xipm_analytic}
Measures that are more convenient for the inference of cosmological information from observational data are the shear correlation functions $\xi_\pm$, which are defined as \begin{equation}
\xi_\pm(\theta) = \la \gamma_{\rm t}\gamma_{\rm t}\ra(\theta) \pm \la \gamma_\times\gamma_\times\ra(\theta) \, .
\end{equation}
Here, $\gamma_{\rm{t}}$ and $\gamma_\times$, denote the tangential and cross-component of the shear for a galaxy pair with respect to their relative orientation \citep[see][]{2002A&A...396....1S}.
The shear correlation functions are the prime estimators to quantify a cosmic-shear signal, since it is simple to include a weighting of the shear measurements into the correlation functions and, contrary to the power spectrum, one does not have to worry about the shape of the survey footprint or masked regions, or model the noise contribution. 
For this analysis, we follow the assumption that a deeper pointing shows a stronger shear signal $\gamma^{\rm{obs}}(\b\theta) = W(\b\theta)\gamma(\b\theta)$ as described above. This assumption implies that a higher redshift just increases the amplitude of the shear signal, but as can clearly be seen by inspecting shear correlation functions of different redshift distributions, the change of the signal is extremely scale-dependent and not just a multiplication with a constant factor. In other words, not just the average shear changes as a function of redshift, but also its entire two-point statistics. However, this should serve as a reasonable first approximation for small variations in mean source redshift. Additionally, we assume that a greater depth does not only lead to a stronger average shear, but also to a higher galaxy number density, implying a correlation between those two quantities.

We denote by $N^i(\b \theta)$ the average weighted number of galaxies\footnote{Instead of using the actual number of galaxies, we take the effective number density, as defined in \citet{Kuijken:2015}, scaled by the respective survey area. Due to this, we account for different weighting of galaxies in the shear correlation functions as well as in the average redshift distribution.} per pointing in redshift bin $i$ and by $W^i(\b \theta)$ the weighting of average shear. The observed correlation functions $\xi^{ij,\text{obs}}_\pm(\theta)$ now change from uniform depth, $\xi_\pm^{ij,\rm{uni}}(\theta)$, via %\citep[see][]{2002A&A...396....1S}
\begin{align}
\xi^{ij,\text{obs}}_\pm(\theta) = & \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_{\rm t}(\bth)\gamma^{j,\rm{obs}}_{\rm t}(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \\
 & \pm \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_\times(\bth)\gamma^{j,\rm{obs}}_\times(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \nonumber\\
 = & \frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i(\bth)N^j(\bthp)\ra} \xi_{\pm}^{ij,\rm{uni}}(\theta) \, ,
 \label{eq:xipmblub1}
 \end{align}
 where the average, $\la\cdot\ra$, represents both an ensemble average as well as an average over the position $\bth$.
 Assuming that the depth of different pointings is uncorrelated, the only important property of a galaxy pair is whether or not they lie in the same pointing. We denote the probability that a random galaxy pair of separation $\theta$ lies in the same pointing by $E(\theta)$. This function is depicted in Fig. \ref{fig:eoftheta_lin}, and an analytic expression is derived in App. \ref{sec:model_e}.

@@

 Ignoring that the depth is correlated on scales of one pointing (here at $\theta \leq \sqrt{2}^\circ$) is equivalent to setting $E(\theta)\equiv 0$. We note that there is still a correlation between $N$ and $W$ for the same galaxy.

@@

 Performing the same calculations as above, this yields a relation between the correlation function of uniform depth, $\xi_\pm^{ij,\rm{uni}}$, and the one that is usually modeled, $\xi_\pm^{ij}$:

@@

When an observer now calculates the model correlation functions $\xi^{ij}_\pm$ without accounting for varying depth between pointings, the ratio between modeled and observed correlation functions becomes: \begin{align}

@@

It is interesting to note that $\xi^{ij}_\pm = \xi_\pm^{ij,\rm{obs}}$ holds wherever $E(\theta)=0$, so we expect the observed and the modeled correlation functions to be equivalent on scales where the depth is uncorrelated. One thing left to determine is how to define the weight-function $W(\b\theta)$. For this, we will refer the reader to the beginning of Sec.~\ref{sec:results}.

@@

\section{A semi-analytic model}
\label{sec:xipm_semianalytic}
The previously derived analytic model describes, how varying depth between pointings modifies the correlation function due to the correlation between number density and the average redshift of source galaxies. While this model serves as an intuitive first approximation, it completely ignores any effects from the large scale structure (LSS) between the closest and the most distant galaxy. Therefore, we do not expect this model to yield accurate, quantitative results for cosmic shear surveys.

@@

An inspection of KiDS-data showed that the redshift distribution of sources is highly correlated with the limiting magnitude in the $r$-band. We thus chose to separate the survey into ten quantiles, sorted by $r$-band depth, that is, if a pointing had a shallower depth than 90\% of the other pointings, it would belong to the first quantile, and so on. For each quantile $m$ and each tomographic redshift bin $i$ we can extract a weighted number of galaxies $N^i_m$ and a source redshift distribution $p^i_m(z)$ following the direct spectroscopic calibration method of H18. In Fig.~\ref{fig:nz_of_meanz}, the average redshift and weighted number of galaxies are plotted for each quantile of each redshift bin, whereas a selection of source redshift distributions is depicted in Fig.~\ref{fig:nofz_quantiles}. A table of the limiting magnitudes for each quantile can be found in Tab.~\ref{tab:maglim}.
%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{images/cov_nz_meanz.pdf}
\caption{Weighted number of galaxies $N$ and average redshift $\la z\ra$ in the KiDS+VIKING-450 survey \citep[KV450,][]{Wright:2018} in pointings of different depth for each of the five tomographic bins used in H18. Each color corresponds to one redshift bin of H18. A single point represents one quantile of the respective redshift bin, where the fainter points denote pointings of shallower depth.}
\label{fig:nz_of_meanz}
\end{figure}

\begin{figure}
\centering
\includegraphics[trim = {1cm 0 1.2cm 0.1cm}, clip, width=\linewidth]{images/dist_z_all.pdf}
\caption{Source redshift distributions $p^i_m(z)$ for a selection of very shallow pointings (blue), average pointings (yellow) and very deep pointings (green). The percentage points in the legend denote to which quantile a pointing belongs, when all are ordered by their depth.}
\label{fig:nofz_quantiles}
\end{figure}
% Asserting an arbitrary cosmological model, we can transform the source redshift distributions $p^i_m(z)$ into comoving distance probability distributions $\mathcal{L}^i_m(\chi)$.

Given the two comoving distance probability distributions of sources, $\mathcal{L}^i_m(\chi)$ and $\mathcal{L}^j_n(\chi)$, we can compute the shear correlation functions from the underlying matter power spectrum, $P_\delta(k,\chi)$, via \citep{1992ApJ...388..272K} \begin{align}
\label{eq:xipm-pkappa}
\xi_{\pm,mn}^{ij}(\theta) =& \int_0^\infty \frac{{\rm d}\ell\,\ell}{2\pi}\, J_{0,4}(\ell\theta)\, P^{ij}_{\kappa,mn}(\ell)\, , \\
\label{eq:pkappa-pdelta/lenseff}
P^{ij}_{\kappa,mn}(\ell) =& \frac{9 H_0^4\Omega_{\rm m}^2}{4c^4}\int_0^{\chi_{\rm{H}}} {\rm d}\chi\, \frac{g^i_m(\chi)g^j_n(\chi)}{a^2(\chi)}\, P_\delta\left(\frac{\ell}{f_K(\chi)},\chi\right)\, , \\
\label{eq:lenseff}
g^i_m(\chi) =& \int_\chi^{\chi_{\rm{H}}} {\rm d}\chi' \, \mathcal{L}^i_m(\chi') \, \frac{f_K(\chi'-\chi)}{f_K(\chi')}\, .
\end{align}
Here, $J_n$ denotes the $n$-th order Bessel Functions, $f_K(\chi)$ is the comoving angular diameter distance and $\chi_{\rm{H}}$ is the comoving distance to the horizon. The parameters $H_0$ and $c$ denote the Hubble constant and the speed of light.
